Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2017-07-22 Response.tex

(In the revised manuscript, the citation numbers for equations, assumptions, propositions, and references are changed. This answer is written according to the new item numbers.)

\subsection*{Reviewer 1} 
\begin{itshape}
Reviewer \#1: I have some issues with this paper that, in my opinion necessitate a major revision.
These issues are:
\end{itshape}

\begin{itemize}
    \item 
        \begin{itshape}
-It extensively re-covers issues that have been covered in many papers, making it difficult to disseminate what the original contributions are.  For example, I don't think the addition of the thrust term into the dynamics necessitates re-doing every piece of analysis that has been done with various resources in the literature about variational integrators (including comparisons with RK methods).  Things that can be found in other research should be cited, and summary comments can be used to point the readers to fill in this knowledge.  The content of this paper easily be reduced considerably to maintain focus on the original contributions.
    \end{itshape}

    Thank you for the comment, and we agree that there is significant repetition of material previously covered in other sources.
    The manuscript has been reduced and reorganized to enhance the clarity and clearly seperate previously material and the contributions of this paper.
    The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
    The second section has been retitiled to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, followed by conclusions.
The section on variational integrators has been reduced and now only discusses the main developments of the integrator for the PCRTBP and omits the review of variational principles.

@@

-To produce constant thrust magnitude or no thrust magnitude (typical of nuclear electric propulsion, or very simple electric propulsion systems.)

-To produce variable thrust magnitude associated with various throttle points at different power accessibility levels (solar electric propulsion).
That is not to say that useful solutions cannot be extracted from the problem, as the authors have shown.  However, a great many solutions in this approach will be unusable due to the fact that acceleration and thrust magnitude are different, i.e. a=T/m

Thus - a MAJOR focus of future work should be the introduction of a seventh state (mass).  This gives rise to a more challenging optimal control problem requiring the switching function, but it is the form that is generally needed in computing optimal open-loop trajectories that can actually be flown.
\end{itshape}

Thank you for the comment and more focus has been included on the pros and cons of this proposed method.
Additional detail has been added to the conclusion which discusses the advantages and the disadvantages, as well as the avenues for future research.
The conclusion is copied below.
\begin{correction}
In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
Direct optimal control techniques must rely entirely on the ability of the numerical optimzation routine to determine a feasible solution.
Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
Furthermore, our approach relies heaviliy on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.

There is additional research to extend these results to more general transfer scenarios in future work.
The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
This will result in a more complicated optimal control problem and is a focus of future research.
Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
\end{correction}

@@

\item 
    \begin{itshape}
The pros and cons of the PCRTBP

-The PCRTB gives rise to exploiting Poincare sections in a systematic manner.  2D solutions might be close enough for qualitative understanding of the solution space for many problems (as long as large plane changes aren't needed).

-The PCRTB is limiting for many pragmatic departure orbits (e.g. LEO and GTO, the most common departure orbits for low-thrust transfers).
\end{itshape}

Thank you for the comment and these characteristics of our choice of the PCRTBP.
Additional detail has been included in the problem formulation section.
This includes some discussion on the choice of the PCRTBP and the relative strengths/weaknesses of this approach.
The additional paragraph has been copied below.

@@

\item 
    \begin{itshape}
The pros and cons of variational integrators should really be acknowledged rather than completely dismissing the RK methods which are still the current state of the art in mission design (using higher-orders and suitable tolerances)

-Pros = boundedness in the energy behavior.  Very long term integration will not deviate from the known dynamical structure.

-Cons = you don't know where you are on the energy surface, just that you are on it somewhere.  This is still a problem in actual trajectory design.  Knowing the error is on an energy surface does not mean that you don't have error.
\end{itshape}

Thank you for the comment and the detail on the benefits/drawbacks of variatonal integrators.
Additional discussion has been included in the paper in the subsection which introduces the variational integrator for the PCRTBP.
Much of the discussion of variational integrators in the introduction has also been removed.
The additional content, which discusses the relative benefits of variational integrators is copied below.

\begin{correction}
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the convential integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The fixed step size of the presented variational integrator mitigates the artifical ``noise'' of variable step integration schemes in gradient computation.
\end{correction}
\end{itemize}
From here on out comments specific to various sections of the paper follow:

### 2017-07-24 manuscript.tex
Correct some typos

\maketitle{}        
\begin{abstract}
We develop a computational approach for the design of continuous low thrust transfers in the planar circular restricted three-body problem.
The use of low thrust propulsion allows the spacecraft to depart from the natural dynamics and enables a wider range of transfers.
We generate the reachable set of the spacecraft and use this to determine transfer opportunities, analogous to the intersection of control-free invariant manifolds.
The reachable set is developed on a lower dimensional \Poincare section and used to design transfer trajectories. 
%Computational geometric mechanics allows for the development of variational integrators and a discrete optimal control problem.
This is solved numerically as a discrete optimal control problem using a variational integrator, which preserve the geometric structure of the motion in the three-body problem.
We demonstrate our approach with two numerical simulations of transfers in the Earth-Moon three-body system.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}. 
Inspite of the long history in astrodynamics, a number of difficulties arise when implenting optimal control techniques.
First, the optimal control problem is typically solved using direct methods, which approximate the continuous problem as a parameter optimization problem.
Alternatively, indirect methods apply the calculus of variations to derive the necessary conditions for optimality.
The indirect approach results in a lower dimensioned problem than the direct approach and provides algebraic conditions that guarantee local optimality in contrast to direct methods which generally rely on suboptimal representations of the control.
Second, in order to implement any optimization method a sufficiently accurate ``initial guess'' is required.
Frequently, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system.
Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
As a result numerical methods are often required to investigate solutions to the three body problem.
Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g. on the order of decades to centuries, which is typical of the three-body problem.

% control-free trajectories in teh three-body problem 
There exists a rich structure in the three-body problem~\cite{koon2011,ross2006}.
Within the three-body problem, a spacecrafts feasible region of motion is constrained by its energy, or Jacobi integral.
It has been shown that there exist multi-dimensional structures, termed invariant manifolds, of constant energy trajectories that span the state space. 
These invariant manifolds, which are associated with the periodic solutions of the three-body dynamics, allow for the spacecraft to traverse vast expanses of the state space with zero energy change. 
However, the results presented are highly case specific and difficult to generalize to arbitrary transfers.
Also, these results are based on control-free trajectories which rely on the underlying structure of the three-body system. 
The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from teh free motion trajectory to allow for increased transfer opportunities.

@@

We defne an optimal control problem to determine the reachability set on a \Poincare~section.
The reachable set, defined as the set of all attainable states subect to the system constraints, allows for the extension of the previous control-free methods based on invariant manifolds.
Defining the reachable set on a \Poincare~section reduces the dimensionality of the system dynamics to the study of a related discrete update map.
While the \Poincare~section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
Instead of relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on the \Poincare~section.
Through the use of low-thrust control input, the reachable set on the \Poincare~section is enlarged and enables a larger space of potential transfers.
By iteratively computing the reachable set, and minimizing the distance to the target on the \Poincare~section allows for the creation of general transfer trajectories.

@@

% discuss some of the pros/cons of using PCRTBP model
We utilize the planar circular restricted three body problem~(PCRTBP) as the basis of our system definition. 
It is a popular model in the preliminary analysis of multibody spacecraft trajectories. 
In the context of Earth based mission, the PCRTBP affords a relatively simple dynamic model while still capturing the major third body perturbation of the Moon.
In addition, this model allows for a systematic process to define and exploit \Poincare sections.
Furthermore, the 2D solutions afforded by the PCRTBP are frequently use to gain a qualitative understanding of the trade space of transfers in the Earth-Moon system.
The PCRTBP approach offers insight into the fundamental dynamical structure while capturing the major dynamic properties of the planar motion.
As a result, this approach is best suited for trajectories which do not require large plane changes.

@@

A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
    The variational integrator uses a range of step sizes between \SIrange{4.72}{472}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
    The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the convential integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The fixed step size of the presented variational integrator mitigates the artifical ``noise'' of variable step integration schemes in gradient computation.

@@

In~\cref{fig:multiple shooting}, we show a schematic representation of the multiple shooting procedure.
We split the optimal control horizon into equal length subsegments such that the length of each segment is \( \frac{k}{n} \), where \( k, n \) are the total number of steps and number of stages, respectively.
Similarly, we divide the state and costate trajectories into \( n \) equal segments. 
To ensure continuity, additional interior constraints are incorporated as
\begin{subequations}
\begin{align}\label{eq:interior_constraints}
        \vecbf{x}_1^{-} - \vecbf{x}_1^{+} &= 0 , \\ 
        \vecbf{\lambda}_1^{-} - \vecbf{\lambda}_1^{+} &= 0 , \\
        \vecbf{x}_2^{-} - \vecbf{x}_2^{+} &= 0 , \\ 
        \vecbf{\lambda}_2^{-} - \vecbf{\lambda}_2^{+} &= 0 , \\
        &\vdots \\
        \vecbf{x}_{n-1}^{-} - \vecbf{x}_{n-1}^{+} &= 0 , \\ 
        \vecbf{\lambda}_{n-1}^{-} - \vecbf{\lambda}_{n-1}^{+} &= 0.
\end{align}
\end{subequations}
Using the multiple shooting method reduces the sensitivity of the terminal states, \( \vecbf{x}_i^{+}, \vecbf{\lambda}_i^{+}\), to variations of the initial states, \(\vecbf{x}_{i-1}^{-}, \vecbf{\lambda}_{i-1}^{-}\).
As a result, the design vector \( \vecbf{\chi} \) is augmented with the additional interior initial conditions, \( \vecbf{x}_i^{-}, \vecbf{\lambda}^{-}\).
Similarly, the constraint vector is augmented with the additional interior constraints defined in~\cref{eq:interior_constraints}.
Based on experimentation, we use four stages in our multiple shooting method.
This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
The multiple shooting algorithm now varies the design vector \( \vecbf{\chi}\) to ensure that the constraints in \( \vecbf{G}(\vecbf{\chi}) \) are satisfied.
In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
% tpbvp is solved using a shooting method (fsolve in matlab)
% determine a design vector X of variables we need to determine
% define a constraint vector G(X) that we need to satisfy
% multiple shooting has interior points of continuity constraints
% sensitivity of terminal state is used to correct the initial condition (cite a book/paper)
% plot of multiple shooting ( stages)
\section{Numerical Example}\label{sec:simulation}
We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
These simulations enable the spacecraft to depart from the natural dynamics through the use of low-thrust propulsion.
The reachability set on the \Poincare section allows for a straightforward method of determining transfer opportunities.
The first example is a transfer from a periodic orbit of the \( L_1 \) Lagrange point to a fixed orbit of the moon.
This example uses a single iteration of the reachable set computation in the design of the transfer.
The second example is a transfer from geostationary orbit of the Earth to a period orbit of \( L_1 \). 
This examples demonstrates the ability to extend the reachability process to multiple iterations, to allow for a much larger and more general transfer.
With both examples it is possible to depart from the vicinity of the Earth to a Moon orbit via a series of reachable sets defined on \Poincare~sections.

@@

% reachable set transfer
Next, we determine the reachability set with the addition of a low thrust control input.
We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of \SI{1}{\newton}.
Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit.

@@

We first generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
The stable manifold is propagated to the \Poincare section at \( y = 0 \), which is denoted by the horizontal black line in the position space visualizations given in~\cref{fig:stage1to4_reachability,fig:stage5to8_reachability,fig:geo_transfer_full}.
The stable manifold is denoted by the green trajecotries which span the majority of the phase space. 
In addition, we plot the intersection of the stable manifold with the \Poincare section with green markers.
The initial orbit and the stable invariant manifold do not intersect and the transfer objective is to enlarge the reachability set to include the stable manifold.

@@

    \begin{subfigure}[htbp]{0.5\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
        \caption{Stage 4 Trajectory~\label{fig:stage4_trajecotry_zoom}} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.5\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage4_poincare.pdf} 
        \caption{Stage 4 \Poincare~section \label{fig:stage4_poincare}} 
    \end{subfigure}   
    \caption{Stage 1-4 reachability sets: The first four reachability sets visualized in both the position (left) and \Poincare space (right).
        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit~\label{fig:stage1to4_reachability}}
\end{figure}
The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_{max} = 0.75 \).
The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) in~\cref{eq:constraints} defined on the \Poincare section.
This allows us to approximate the set of states that are achievable in the \( \parenth{x ,\dot{x}} \) space.
The intersection of the first reachable set with the \Poincare section is shown in~\cref{fig:stage1_poincare} in red.
While the reachable set does not intersect the stable manifold, it does reduce some of the distance in the \(\dot{x}\) dimension.
From this reachable set, we chose a trajectory which minimizes the distance towards the stable manifold, which is shown in~\cref{fig:stage1_poincare} by the black marker.
The minimum trajectory is also shown in~\cref{fig:stage1_trajecotry_zoom} and is used to initilize the following stage.
Over the relatively short time horizon of the geostationary orbital period, the reachability set remains quite close to the initial orbit. 
However, the addition of the control input has increased the velocity component of the trajectory.
It is this velocity change that is subsequently exploited to generate the transfer trajectory.

@@

After each computation of the reachability set we pick a state which minimizes the distance on the \Poincare section towards the stable manifolds as the function \( d \)
\begin{align*}
        d(\vecbf{x}(N)) = \sqrt{\parenth{x - x_t}^2 + \parenth{\dot{x} - \dot{x}_t}^2}.
\end{align*}
From the reachable set, the trajectory which minimizes \( d \) is used to intialize the next iteration.
Once the reachability set intersects the stable manifold, we can achieve a complete transfer from the initial geostationary orbit to the stable manifold of \( L_1 \).

>>>>

\maketitle{}        
\begin{abstract}
We develop a computational approach for the design of continuous low thrust transfers in the planar circular restricted three-body problem.
The use of low thrust propulsion allows the spacecraft to depart from the natural dynamics and enables a wider range of transfers.
We generate the reachable set of the spacecraft and use this to determine transfer opportunities, analogous to the intersection of control-free invariant manifolds.
The reachable set is developed on a lower dimensional \Poincare section and used to design transfer trajectories. 
%Computational geometric mechanics allows for the development of variational integrators and a discrete optimal control problem.
This is solved numerically as a discrete optimal control problem using a variational integrator, which preserves the geometric structure of the motion in the three-body problem.
We demonstrate our approach with two numerical simulations of transfers in the Earth-Moon three-body system.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

There has been extensive research focused on the optimal control of spacecraft orbital transfers in the three-body problem~\cite{mingotti2011,grebow2011}. 
Inspite of the long history in astrodynamics, a number of difficulties arise when implenting optimal control techniques.
First, the optimal control problem is typically solved using direct methods, which approximate the continuous problem as a parameter optimization problem.
Alternatively, indirect methods apply the calculus of variations to derive the necessary conditions for optimality.
The indirect approach results in a lower dimensioned problem than the direct approach and provides algebraic conditions that guarantee local optimality in contrast to direct methods which generally rely on suboptimal representations of the control.
Second, in order to implement any optimization method a sufficiently accurate ``initial guess'' is required.
Frequently, insight into the problem or intuition on the part of the designer is often required to determine initial conditions that will converge to the desired solution due to the inherent nonlinear and chaotic behavior of the three-body system.
Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
As a result, numerical methods are often required to investigate solutions to the three body problem.
Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
Therefore, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g. on the order of decades to centuries, which is typical of the three-body problem.

% control-free trajectories in teh three-body problem 
There exists a rich structure in the three-body problem~\cite{koon2011,ross2006}.
Within the three-body problem, a spacecrafts feasible region of motion is constrained by its energy, or Jacobi integral.
It has been shown that there exist multi-dimensional structures, termed invariant manifolds, of constant energy trajectories that span the state space. 
These invariant manifolds, which are associated with the periodic solutions of the three-body dynamics, allow for the spacecraft to traverse vast expanses of the state space with zero energy change. 
However, the results presented are highly case specific and difficult to generalize to arbitrary transfers.
Also, these results are based on control-free trajectories which rely on the underlying structure of the three-body system. 
The addition of low-thrust propulsion offers the potential of reduced transfer transit times and the ability to depart from the free motion trajectory to allow for increased transfer opportunities.

@@

We defne an optimal control problem to determine the reachability set on a \Poincare~section.
The reachable set, defined as the set of all attainable states subject to the system constraints, allows for the extension of the previous control-free methods based on invariant manifolds.
Defining the reachable set on a \Poincare~section reduces the dimensionality of the system dynamics to the study of a related discrete update map.
While the \Poincare~section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
Instead of relying on intuition or insight into the problem, states are chosen which minimize the distance towards the target on the \Poincare~section.
Through the use of low-thrust control input, the reachable set on the \Poincare~section is enlarged and enables a larger space of potential transfers.
By iteratively computing the reachable set, and minimizing the distance to the target on the \Poincare~section, we generate general transfer trajectories.

@@

% discuss some of the pros/cons of using PCRTBP model
We utilize the planar circular restricted three body problem~(PCRTBP) as the basis of our system definition. 
It is a popular model in the preliminary analysis of multibody spacecraft trajectories. 
In the context of Earth based mission, the PCRTBP affords a relatively simple dynamic model while still capturing the major third body perturbation of the Moon.
In addition, this model allows for a systematic process to define and exploit \Poincare sections.
Furthermore, the 2D solutions afforded by the PCRTBP are frequently used to gain a qualitative understanding of the trade space of transfers in the Earth-Moon system.
The PCRTBP approach offers insight into the fundamental dynamical structure while capturing the major dynamic properties of the planar motion.
As a result, this approach is best suited for preliminary trajectories which do not require large plane changes.

@@

A simulation comparing the variational integrator to a conventional Runge-Kutta method is presented in this section.
A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0 & 0.2883\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
    The variational integrator uses a range of step sizes between \SIrange{4.72}{472}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
    The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the convential integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The fixed step size of the presented variational integrator mitigates the artifical ``noise'' of variable step integration schemes in gradient computation.

@@

In~\cref{fig:multiple shooting}, we show a schematic representation of the multiple shooting procedure.
We split the optimal control horizon into equal length subsegments such that the length of each segment is \( \frac{k}{n} \), where \( k, n \) are the total number of steps and number of stages, respectively.
Similarly, we divide the state and costate trajectories into \( n \) equal segments. 
To ensure continuity, additional interior constraints are incorporated as
\begin{subequations}
\begin{align}\label{eq:interior_constraints}
        \vecbf{x}_1^{-} - \vecbf{x}_1^{+} &= 0 , \\ 
        \vecbf{\lambda}_1^{-} - \vecbf{\lambda}_1^{+} &= 0 , \\
        \vecbf{x}_2^{-} - \vecbf{x}_2^{+} &= 0 , \\ 
        \vecbf{\lambda}_2^{-} - \vecbf{\lambda}_2^{+} &= 0 , \\
        &\vdots \\
        \vecbf{x}_{n-1}^{-} - \vecbf{x}_{n-1}^{+} &= 0 , \\ 
        \vecbf{\lambda}_{n-1}^{-} - \vecbf{\lambda}_{n-1}^{+} &= 0.
\end{align}
\end{subequations}
Using the multiple shooting method reduces the sensitivity of the terminal states, \( \vecbf{x}_i^{+}, \vecbf{\lambda}_i^{+}\), to variations of the initial states, \(\vecbf{x}_{i-1}^{-}, \vecbf{\lambda}_{i-1}^{-}\).
As a result, the design vector \( \vecbf{\chi} \) is augmented with the additional interior initial conditions, \( \vecbf{x}_i^{-}, \vecbf{\lambda}^{-}\).
Similarly, the constraint vector is augmented with the additional interior constraints defined in~\cref{eq:interior_constraints}.
Based on experimentation, we use four stages in our multiple shooting method.
This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
The multiple shooting algorithm now varies the design vector \( \vecbf{\chi}\) to ensure that the constraints in \( \vecbf{G}(\vecbf{\chi}) \) are satisfied.
In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
% tpbvp is solved using a shooting method (fsolve in matlab)
% determine a design vector X of variables we need to determine
% define a constraint vector G(X) that we need to satisfy
% multiple shooting has interior points of continuity constraints
% sensitivity of terminal state is used to correct the initial condition (cite a book/paper)
% plot of multiple shooting ( stages)
\section{Numerical Example}\label{sec:simulation}
We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
These simulations enable the spacecraft to depart from the natural dynamics through the use of low-thrust propulsion.
The reachability set on the \Poincare section allows for a straightforward method of determining transfer opportunities.
The first example is a transfer from a periodic orbit of the \( L_1 \) Lagrange point to a fixed orbit of the moon.
This example uses a single iteration of the reachable set computation in the design of the transfer.
The second example is a transfer from geostationary orbit of the Earth to a period orbit of \( L_1 \). 
This examples demonstrates the ability to extend the reachability process to multiple iterations, to allow for a much larger and more general transfer.
With both examples it is possible to depart from the vicinity of the Earth to a Moon orbit via a series of reachable sets defined on \Poincare sections.

@@

% reachable set transfer
Next, we determine the reachability set with the addition of a low thrust control input.
We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of approximately \SI{1}{\newton}.
Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit.

@@

We first generate the stable invariant manifold associated with the periodic orbit in order to determine our target set.
The stable manifold is propagated to the \Poincare section at \( y = 0 \), which is denoted by the horizontal black line in the position space visualizations given in~\cref{fig:stage1to4_reachability,fig:stage5to8_reachability,fig:geo_transfer_full}.
The stable manifold is denoted by the green trajectories which span the majority of the phase space. 
In addition, we plot the intersection of the stable manifold with the \Poincare section with green markers.
The initial orbit and the stable invariant manifold do not intersect and the transfer objective is to enlarge the reachability set to include the stable manifold.

@@

    \begin{subfigure}[htbp]{0.5\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
        \caption{Stage 4 Trajectory~\label{fig:stage4_trajecotry_zoom}} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.5\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage4_poincare.pdf} 
        \caption{Stage 4 \Poincare~section \label{fig:stage4_poincare}} 
    \end{subfigure}   
    \caption{Stage 1-4 reachability sets: The first four reachability sets visualized in both the position (left) and \Poincare space (right).
        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit~\label{fig:stage1to4_reachability}}
\end{figure}
The first reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section and we again assume a upper bound on the thrust magnitude of \( u_{max} = 0.75 \).
The reachable set is generated by varying the angle \( \ang{0} \leq \theta < \ang{360} \) in~\cref{eq:constraints} defined on the \Poincare section.
This allows us to approximate the set of states that are achievable in the \( \parenth{x ,\dot{x}} \) space.
The intersection of the first reachable set with the \Poincare section is shown in~\cref{fig:stage1_poincare} in red.
While the reachable set does not intersect the stable manifold, it does reduce some of the distance in the \(\dot{x}\) dimension.
From this reachable set, we chose a trajectory which minimizes the distance towards the stable manifold, which is shown in~\cref{fig:stage1_poincare} by the black marker.
The distance on the \Poincare section between the reachable set and the stable manifolds is defined as the function \( d \),
\begin{align*}
        d(\vecbf{x}(N)) = \sqrt{\parenth{x - x_t}^2 + \parenth{\dot{x} - \dot{x}_t}^2}.
\end{align*}
From the reachable set, the trajectory which minimizes \( d \) is used to intialize the next iteration.
The minimum trajectory is also shown in~\cref{fig:stage1_trajecotry_zoom} and is used to initilize the following stage.
Over the relatively short time horizon of the geostationary orbital period, the reachability set remains quite close to the initial orbit. 
However, the addition of the control input has increased the velocity component of the trajectory.
It is this velocity change that is subsequently exploited to generate the transfer trajectory.

@@
